---
title: "data-analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{data-analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  error = FALSE,
  message = FALSE)
```

```{r}
library(stakeholderAnalysis)
library(tidyverse)
library(readxl)
library(pheatmap)
library(tidymodels)
```

## Data

### Questionnaire

```{r}
data_stakeholders_clean <- stakeholderAnalysis::data_stakeholders_clean
```

### Clusters

```{r}
data_clusters <- stakeholderAnalysis::data_clusters

```


## Select numeric columns

```{r}
map(
  data_stakeholders_clean,
  safely(as.numeric)
  ) |> transpose() |>
  pluck("result") |>
  as_tibble() -> data_numeric

## text columns
col_ind <- map_df(
  data_stakeholders_clean,
  as.integer
)

not_any_na <- function(x) all(!is.na(x))

data_text_ind <- map_lgl(
  col_ind,
 not_any_na
)

data_text <- data_stakeholders_clean[ , !data_text_ind]

## remove all NA columns
not_any_na <- function(x) all(!is.na(x))

data_numeric <- data_numeric |>
  select(where(not_any_na))

## convert all to counts
#map_df(
#  data_numeric,
#  as.numeric
#)

## add name of org
data_numeric$type_org <- data_stakeholders_clean$organis
```

## Principal Component analysis

```{r}
theme_set(theme_bw(16))
```

```{r}
#data_numeric$label <- data_stakeholders_clean$organis

pca_recipe <- recipe(~., data = data_numeric)

pca_trans <- pca_recipe %>%
  # center the data
  step_center(all_numeric()) %>%
  # center the data
  step_scale(all_numeric()) %>%
  # pca on all numeric variables
  step_pca(all_numeric())

pca_estimates <- prep(pca_trans)

# std deviation
sdev <- pca_estimates$steps[[3]]$res$sdev
percent_variation <- sdev^2 / sum(sdev^2)
var_df <- data.frame(PC=paste0("PC",1:length(sdev)),
                     var_explained=percent_variation,
                     stringsAsFactors = FALSE)

var_df %>%
  mutate(PC = fct_inorder(PC)) %>%
  ggplot(aes(x=PC,y=var_explained)) + geom_col()

juice(pca_estimates) %>%
  ggplot(aes(PC1, PC2)) +
  geom_point(aes(color = type_org), size = 5) +
  labs(title="PCA from tidymodels")
```

## Create tidy, long format

```{r}
data_numeric_tidy <- data_numeric |>
  pivot_longer(cols = -c(type_org), values_to = "score", names_to = "question")
```

## Add clusters

```{r}
data_numeric_tidy <- data_numeric_tidy |>
  separate(question, into = c("question", "sub", "trace"), remove = FALSE) |>
  mutate(
    question = map(
      question,
      str_replace_all,
      pattern = "x",
      replacement = ""
      )
    ) 

data_numeric_tidy$question <- data_numeric_tidy$question |> unlist() |> as.integer()

data_clusters$question <- data_clusters$question |>
  as.integer()

data_numeric_tidy_clusters <- left_join(
  data_numeric_tidy, 
  data_clusters
  )

```

## Exploratory plots

```{r}
data_numeric_tidy_clusters |>
  ggplot(
    aes(
      x = question,
      y = score)
  ) + 
  geom_point(aes(colour = type_org), position = "jitter") +
  xlab("Question number") +
  ylab(NULL) + toolboxr::rotate_axis_labels("x", 90)

data_numeric_tidy_clusters |>
  ggplot(
    aes(
      x = question,
      y = score)
  ) + 
  geom_point(aes(colour = type_org), position = "jitter") +
  xlab("Question number") +
  ylab(NULL) + toolboxr::rotate_axis_labels("x", 90) +
  facet_wrap(~cluster_name)

```

## Aggregate organization types

For some organization types there are only a few observations. Let's aggregate the organization types into 4 subtypes:

-   Government & Policy
-   Industry
-   NGO
-   Academia

```{r}
names(data_numeric_tidy_clusters)

data_numeric_tidy_clusters |> 
  group_by(type_org) |>
  tally()


data_numeric_tidy_clusters <- data_numeric_tidy_clusters |>
  mutate(
  type_collapsed = fct_collapse(type_org,
  `Government & Policy` = c("EU Commision", "EU agency", "Regulatory authority"),
  Academia = c("Policymakers - academia"),
))


data_numeric_tidy_clusters |> 
  group_by(type_collapsed) |>
  tally()

```

## Polar plots
```{r}

data_numeric_tidy_clusters |>
  group_by(question, cluster_name, cluster_id) |>
  summarise(
    median_score = median(score, na.rm = TRUE),
    sd = sd(score, na.rm = TRUE)) |>
  ggplot(
    aes(
      x = reorder(as_factor(question), cluster_id),
      y = median_score)
  ) + 
  geom_bar(
    aes(
      fill = cluster_name), 
    stat="identity", 
    color="black", 
    position = "dodge") +
  geom_errorbar(
    aes(ymin=median_score-sd, 
        ymax=median_score+sd),
    width=.2,
    position=position_dodge(.9)) + 
  xlab("Question number") +
  ylab(NULL) + 
  toolboxr::rotate_axis_labels("x", 90) +
  ggtitle("Overall average score per question category")  -> p

p

p + 
  coord_polar() +
  theme_minimal() +
  # Remove legend, axes, text, and tick marks
  theme(
   # legend.position = "none",
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_text(face = "bold"),
    plot.title = element_text(size = 24, face = "bold"),
    plot.subtitle = element_text(size = 12) 
  ) -> pp

pp
```


## Polar plot per organization category

```{r}
data_nested <- data_numeric_tidy_clusters |>
  group_by(type_collapsed) |>
  nest()


data_nested <- data_nested |>
  mutate(polar_org = map2(
    data,
    type_collapsed,
    plot_polar,
    show.legend = FALSE
  ))

data_nested$polar_org
```

```{r, out.width=800}
## panel
cowplot::plot_grid(
  data_nested$polar_org[[1]],
  data_nested$polar_org[[2]],
  data_nested$polar_org[[3]],
  data_nested$polar_org[[4]],
  nrow = 2
)


```

## Heatmap - per org_type

```{r, out.width=800}
data_numeric_num <- data_numeric |>
  select(-c(type_org)) |>
  as.matrix() -> data_mtrx

row.names(data_mtrx) <- data_numeric$type_org
pheatmap(data_mtrx)

```

